\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2022-08-13 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound2-TPWRS-01391-2021R1.tex

{\nib \blue{Thank you. Please find our itemized responses to the questions related to numerical simulations.}}

\noindent
{\bf R4.2. Comments by reviewer}:\\
{\em It is stated that ``The Fig 9 shows that the computational time needed to perform the proximal updates, for all $t_k$, are indeed orders-of-magnitude faster than $h$."
This statement doesn't seem to hold, since $h=10^{-3}$ and the computational time lies within the range of around $10^{-3}$ to $10^{-2}$. How does the computational time needed to perform the proximal updates relate to time step $h$? If they are not strongly related, does that mean the step size $h$ is constrained to be greater than that computational time?}

{\nib \blue{The reviewer is right, we should have been more careful in phrasing that sentence. We meant to say that the computational time needed to perform a proximal update is \ul{statistically} comparable (in the big O sense) to $h$. The reason why the precise computational time is probabilistic is twofold: random sampling, and randomness in computational latency, power and memory footprint. Fig. 9 shows that $h=10^{-3}$ and the computational time needed to perform a single proximal update is $O(10^{-3})$ with high probability in the sense most time-steps incur $O(10^{-3})$ time. In Fig. 9, while instances when computational times are $O(10^{-2})$ do occur for Case I, they are rare. \ul{In the present revision, we have now rephrased the corresponding text (highlighted in blue) toward the end of Sec. V-A}.

The proximal updates involve certain block coordinate ascent recursions in the dual co-ordinates $\bm{y},\bm{z}$; see lines 13-21 in Algorithm 1. This recursion is guaranteed to be contractive (hence existence, uniqueness and convergence are guaranteed) in the product orthant cone: $(\bm{y},\bm{z})\in\mathbb{R}^{N}_{>0}\times \mathbb{R}^{N}_{>0}$, see ref. [41, Sec. III-B, III-C] and especially Corollary 4 in that reference. We have cited that reference in this manuscript's Sec. IV. \ul{While $h$ appears as parameter in this recursion (lines 5 and 14 in Algorithm 1), we do not have analytical relation between the computational time and $h$}. We did not constrain or adapt $h$ in Algorithm 1. We simply set a small value $h=10^{-3}$ to promote statistical consistency (see the unnumbered equation after (27)). Depending on the random sampling, dynamics and computational platform, it is possible for the single-step proximal update computational time to fall below $h$ (see e.g., ref. [41, Fig. 9] which used the same $h=10^{-3}$). \ul{The standard desktop computer/processor we used for our simulation is reported in Sec. V, first paragraph, last sentence}.}}

@@

\noindent
{\bf R4.3. Comments by reviewer}:\\
{\em In Fig. 10, could the authors please explain why the computational time of Case I is less than that of Case II?
The computational cost for the 5-generator system is roughly 4 times that of Case II. Could the authors comment if such a scale makes sense intuitively considering the state dimension of the two systems?}

{\nib \blue{Thanks for this comment. The state dimensions for both Cases I and II in Fig. 10 are the same, i.e., $5\times2 = 10$ dimensions. Other than the difference in dynamics due to line failure in Case II, we do not think there is any specific physical insight to be gleaned on why the computational times for Case II are below Case I; they both use identical set of random initial samples although the Wiener process noise realizations are different.

Regarding the scaling comment, perhaps the reviewer meant to say that the total computational times for the 50 generator (100 state) system are roughly 4 times that of Case II (5 generator, 10 state). The possible explanation of why the scaling is not so bad is because the number of samples for the 50 generator case is $2000$, which is twice the number of samples used in Cases I and II. This is made possible by our usage of scattered point clouds with explicit proximal computation of the time-varying probability weights, as opposed to gridding the state space and then approximating the joint PDF via histogram (which would have caused exponential in dimension scaling in the computational cost). Being able to directly evaluate the joint PDF values via proximal computation allows us to bypass construction of function approximation oracle over the high dimensional state space (which in general scales poorly with dimensions), and then to evaluate it. A direct comparison between the two cases is less meaningful due to difference in parameter values, initial joint PDF etc. \ul{However, the general trend that the proposed scattered point cloud-type proximal framework, similar to the one proposed here, can incur computational times which scale sublinearly with dimensions while resolving the joint PDF well, were reported in our earlier works in different application contexts: ref. [41, Sec. V-B and Fig. 14], ref. [42, Fig. 5]}.

}}

### 2022-08-13 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev2.tex
Rebuttal round 2

At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1. This proximal update has no analytical solution and instead requires execution of a contractive fixed point recursion. The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update. For the proposed method to be practical, we therefore, need to demonstrate that {\blue{the computational timescale is at least statistically comparable (in the order of magnitude sense) to the physical dynamics timescale. In other words, the computational times needed to complete the zero-order-hold fixed point recursions are, with high probability, the same orders-of-magnitude as $h$, the step-size of physical time. The precise computational time is probabilistic, not just due to random sampling, but also due to randomness in power and memory footprint.}}

@@

{\blue{In Fig \ref{fig:CompTimeIEEE14bus}, $h=10^{-3}$ and the computational time needed to perform a single proximal update, \emph{with high probability}, is $O(10^{-3})$ in the sense most time-steps incur $O(10^{-3})$ time. While instances when computational times are $O(10^{-2})$ do occur for Case I in Fig \ref{fig:CompTimeIEEE14bus}, but are rare.}}

>>>>

At each fixed physical time $t_{k}=kh$, $k\in\mathbb{N}$, where $h$ is the fixed step-size, we perform the proximal update $\varrho_{k-1}\mapsto\varrho_{k}$ via Algorithm 1. This proximal update has no analytical solution and instead requires execution of a contractive fixed point recursion. The physical time $t_{k}$ is ``frozen" (zero-order-hold) during the fixed point recursions underlying this one-step proximal update. For the proposed method to be practical, we therefore, need to demonstrate that {\blue{the computational timescale is at least statistically comparable (in the order of magnitude sense) to the physical dynamics timescale. In other words, the computational times needed to complete the zero-order-hold fixed point recursions are, with high probability, the same orders-of-magnitude as $h$, the step-size of physical time. The precise computational time is probabilistic, not just due to random sampling, but also due to randomness in latency, power and memory footprint of the hardware platform.}}

@@

{\blue{In Fig \ref{fig:CompTimeIEEE14bus}, $h=10^{-3}$ and the computational time needed to perform a single proximal update, \emph{with high probability}, is $O(10^{-3})$ in the sense most time-steps incur $O(10^{-3})$ time. Few instances when computational times are $O(10^{-2})$ do occur for Case I in Fig \ref{fig:CompTimeIEEE14bus}, but are rare.}}
